---
title: "Data Management and Reproducibility"
date: "September 2021" 
author: "Kaleigh Fisher" 
output:
  html_document: default
  pdf_document: default
---

Make change here
hello
# Best practices for data management
1. Organize data and projects in an accessible way - for both the primary investigator and others 
+ Descriptive naming 
+ Structured and consistent file approach 
2. Once you input data - do not change the master file and give it an informative name 
+ I usually try to put the year I input the data and information about the project that distinguishes from other projects in the group I work in. 
+ For example - YearGrantProjectDescription
3. Version control - github 
+ Regularly commit and push code to github
4. Collaborative coding - github 
+ Write code so you can share it at any time during the development process - not just when it is finished
5. Self-contained workflows - 
+everything needed to run the code is accessible
ie. data manipulation and analysis can be executed and critiqued by any person, including the primary investigator
+ Make sure to take lots of notes: describe the data and your code 

## What is github? 
A development platform that has a variety of tools, including collaborative coding. 
-Can create repositories with coding scripts that can be shared and modified by multiple users. 
-Version control 

## Benefits of reproducible data management
1. Remind yourself and others how you organized and analyzed your data - easier to complete manuscript/presentation
2. More transparency, confidence and rigor - part of the peer review process?
3. Greater trust from the scientific community and the general public 
4. Makes science more accessible - more people can understand and critique analyses, not just experts 

For resources on how to set up git desktop, git2r and trouble shooting visit this wiki: https://happygitwithr.com/

# Intro to Github
1. Make a github account
Let's download some data to put into the repository 
This data is from dryad - Do riparian forest strips in modified forest landscapes aid in conserving bat diversity?
https://datadryad.org/stash/dataset/doi:10.5061/dryad.3td78sk
*make sure to save as csv first and then upload to git

# Navigating folders/repositories/directories on the command line (terminal) 
ls: see list of folders 
cd: change directory
mkdir: make a new folder 
mv: move a folder or directory 
example - mv ~/Downloads/doi_10.5061_dryad.k3j9kd56m__v4/input01.csv ~/Desktop/input01.csv

# Intro to Rstudio 
Different windows and features
Kaleigh Change Here

# Github and Rstudio
1.Clone repository onto local desktop via command line in Rstudio or local command line
+ Navigate to folder you want to place repository into (cd) 
To see a list of files in the directory you are currently in: ls; windows: dir
+ In the command line type -> git clone "paste repository name from github"
3. In R-studio (right hand corner), open github repository 
4. Start working :) 
5. Be sure to regularly save and commit your work. At the end of a session, make sure to push back to github. At the beginning of any session, make sure to pull from github, especially if you are working with collaborators. 
-Specifically, after you make a change to the code file (Rmd), do the following in this order. 

1) Save the file with the floppy disk on the top panel or file>save
2) Click on the file that you changed and saved in the git window
3) Press commit in the git window and write the message 
4) Push your changes 

# Branches and Merging 
- Best practice is to make a branch when working on script files rather than the main
1. Make sure your repo is up to date by pulling 
2. click on the shape icon on the upper right window in git tab 
3. Name the branch 
4. commit and push to branch 
5. when ready to merge - go to github and navigate to branches 
6. submit pull request 
7. solve any merge issues  
8. commit changes 

# Now lets look at our data 
```{r}
# First where are we? 

getwd()  # prints out current working directory 

#read.csv("/Users/rawanaljudeid/Desktop/2021RladiesBats/Bats_Capture_Database.csv") # Absolute file path 
data<-read.csv("./Bats_Capture_Database.csv") # Relative file path 

# What are the columns? 
names(data) # can also look at readme file 
# How many samples do we have? 
nrow(data)
print(paste("We have", nrow(data), "bat samples."))

head(data)
ncol(data)
```

# Descriptive Info about Data
```{r}
library(ggplot2) # make pretty graphs
library(ggpubr) # save figures 
library(plyr) # summary statistics 
library(tidyr) # data manipulation
names(data)

# How many bats collected? 
nrow(data)
# How many bat species are in the dataset? 
batspec<-unique(data["Species"]) 

print(paste("There are", nrow(batspec), "species in the dataset."))


# Reduce dataframe
SpeciesRed<-unique(data[c("Habitat", "Species")]) 
SpeciesRed[order(SpeciesRed$Habitat),]
head(SpeciesRed)

# How many species in each habitat type? 
SpeciesRich<-aggregate(Species~Habitat,
                     data=SpeciesRed,
                     FUN=length)
SpeciesRich

# Graph it 
ggplot(data=SpeciesRich, aes(x=Habitat, y=Species)) +
    geom_bar(colour="black", stat="identity")

ggsave("./Figures/BatSpecies.png")

# How many individuals per habitat type and site? 
head(data)

SiteAbund<-aggregate(Forearm~Habitat + Site.Code, 
                     data=data,
                     FUN=length)
# Rename forearm to abundance for clarity 
names(SiteAbund)[names(SiteAbund) == "Forearm"] <- "Abundance"

# What is the average abundance across sites for each habitat type? 
# This generates a dataframe with average and standard deviation so we can plot it
AbundanceSum<-ddply(SiteAbund, c("Habitat"), summarise,
      N = length(Abundance), # Number of sites they collected at for each habitat
      mean = mean(Abundance),
      sd = sd(Abundance),
      se = sd / sqrt(N))
AbundanceSum

ggplot(data=AbundanceSum, aes(x=Habitat, y=mean)) +
    geom_bar(colour="black", stat="identity") +
   theme(axis.text.x  = element_text(angle=90, vjust=0.5, size=16)) + 
  ggtitle("Average Abundance for Each Habitat Type") +
   geom_errorbar(aes(ymin=mean-se, ymax=mean+se),
                  width=.2,                    # Width of the error bars
                  position=position_dodge(.9))
```

# Transform Data and Do Stats with Vegan 
```{r}
library(vegan) # package for diversity statistics 
head(data)
datafilt<-data[, c("Species", "Site.Code", "Habitat", "Forearm")]

datatransform<-aggregate(Forearm~Habitat + Species + Site.Code,
                     data=datafilt,
                     FUN=length)
names(datatransform)
data_vegan<-pivot_wider(datatransform, 
                        names_from=Species, values_from=Forearm)

# Make all NAs 0 so it is compatible with vegan - represents 0 individuals of a species at that site
data_vegan[is.na(data_vegan)]<-0

# Look at data
head(data_vegan)

# remove first two columns and change into a matrix 
dataMatrix<-as.matrix(data_vegan[,3:length(data_vegan)])
dataMatrix
# Now we can run the metaMDS function in vegan to plot an NMDS
nmds_data<- metaMDS(dataMatrix, distance="bray")
stressplot(nmds_data)

# Get NMDS scores
data.scores=as.data.frame(scores(nmds_data))

# merge sample NMDS info with sample info
data.scores$Site<-data_vegan$Site.Code
data.scores$Habitat<-data_vegan$Habitat

head(data.scores)
# Plot it using ggplot
NMDS<-ggplot(data.scores, aes(x=NMDS1, y=NMDS2)) + 
  geom_point(size=4, aes(shape=Site, colour=Habitat))
  
NMDS
ggsave("./Figures/NMDS.png", width=7, height=5)

```




